\documentclass[a4paper]{article}
\usepackage[a4paper, margin=2.56cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage{enumitem}
\lstset
{ %Formatting for code in appendix
    numberstyle=\tiny,
    basicstyle=\ttfamily\footnotesize,
    numbers=left,
    showstringspaces=false,
    breaklines=true,
}

%\title{Sorting of Binary Search Trees in Coq}

\title{%
  Binary Search Trees in Coq \\
  \large Proving with Computer Assistance Assignment}
    
\author{Henrique Dias (1531484), Venislav Varbanov (1284401)}
\date{March 2021}

\begin{document}

\maketitle

\section{Introduction}

The goal of this assignment is to formally define trees and binary search trees of natural numbers, as well as define some of its common operations, such as sort, insert and tree minimum value.

A binary search tree (BST) is a tree that for each of its nodes, its left subtree only stores values smaller than the value of the node, and the right subtree only stores values larger than the value of the node.

Below we have a description of each function and lemma, as well as the code snippets to which they correspond. The lines in the snippets correspond to the \texttt{.v} file annexed in the end.

\section{Part I: Sorting a Tree}

\paragraph{Definition of a BST}

The definition of a BST (\texttt{bst}) is a recursive function where, for each node, we check if its value is greater than all the values in the node's left subtree, and less than the ones in the right subtree. To check those conditions, we defined a helper recursive function \texttt{tree\_forall} (line 17) that for a certain condition and a certain node, checks if all of the nodes in the subtree, rooted at the given node, satisfy the given condition.

\lstinputlisting[firstnumber=16,firstline=16,lastline=36]{bst.v}

\paragraph{Insert Operation}

The insert operation (\texttt{insert}) can be applied to a BST and is recursive by nature.

\lstinputlisting[firstnumber=38,firstline=38,lastline=47]{bst.v}

To prove the correctness of the insertion, we started by proving a sublemma \texttt{insert\_forall}. The sublemma proves that for a certain condition, tree and a value (to be inserted), if all the nodes of the tree meet the condition, then if the given value also meets it, the condition will still be met by all nodes after the insert operation. The proof is by induction and simple.

\lstinputlisting[firstnumber=50,firstline=50,lastline=60]{bst.v}

To prove the correctness of insertion, i.e., to prove that inserting a new value in a BST produces a BST, we made a proof by induction (\texttt{insert\_bst}, line 63) that leverages the sublemma \texttt{insert\_forall}, applying the necessary conditions that a BST needs to meet on its left and right branches.

\lstinputlisting[firstnumber=63,firstline=63,lastline=76]{bst.v}

\paragraph{Sort Operation}

The sort operation (\texttt{sort}) transforms an arbitrary tree into a BST. To do so, we first transform the tree into a list (\texttt{tree\_to\_list}) and then we rebuild a BST from the list (\texttt{list\_to\_bst}) by simply using the insert operation on all list elements.

\lstinputlisting[firstnumber=78,firstline=78,lastline=93]{bst.v}

To prove that the sort operation results in a BST (\texttt{sort\_correct}), we proved a sublemma that shows that any tree built from a list with our insert operation is a BST (\texttt{list\_to\_bst\_correct}). When building a tree from a list, we simply use the insert operation from an empty tree and insert all elements consecutively. Since the \texttt{insert} operation always results in a BST, it follows that \texttt{sort} results in a BST.

\lstinputlisting[firstnumber=95,firstline=95,lastline=108]{bst.v}

We also prove that the sorted tree has the exact same elements as the initial tree (\texttt{sorted\_occurs}). To prove this, we first defined the function \texttt{occurs} that checks if a certain value is present in a tree.

\lstinputlisting[firstnumber=111,firstline=111,lastline=115]{bst.v}

Then, we defined three sublemmas:

\begin{itemize}
    \item \texttt{occurs\_insert}, which proves that, for a tree and two numbers $n$ and $m$ (to be inserted in the tree), if $m=n$ or $n$ occurs in the tree already, then $n$ will occur in the tree after inserting $m$. The check for $m=n$ is needed for the edge case where the initial tree is a leaf. In that case, it is impossible for a number to occur on it.
    \lstinputlisting[firstnumber=118,firstline=118,lastline=133]{bst.v}

    \item \texttt{occurs\_tree\_list}, which proves that if a number occurs in a tree, then it will also occur in its list transformation, and vice-versa.
    \lstinputlisting[firstnumber=136,firstline=136,lastline=152]{bst.v}
    
    \item \texttt{occurs\_list\_tree} (line 155), which proves that if a list contains a number, then it will occur in the BST built from that list, and vice-versa. Here is where we need to apply the \texttt{occurs\_insert} sublemma, as the BST transformation is the one that uses the insert operation.
    \lstinputlisting[firstnumber=155,firstline=155,lastline=165]{bst.v}
\end{itemize}

Using this sublemmas, it is trivial to prove \texttt{sorted\_occurs} by induction.

\lstinputlisting[firstnumber=168,firstline=168,lastline=181]{bst.v}

\section{Part II: The Minimum of a BST}

\paragraph{Tree Minimum} The tree minimum function (\texttt{treeMin}) is defined recursively using the builtin \texttt{Nat.min} function. It returns an \texttt{option nat}, representing the minimum value in the tree, which is the minimum value among the one of the left subtree, the right subtree, and the value of the root. In case the tree is a leaf, \texttt{None} is returned.

\lstinputlisting[firstnumber=186,firstline=186,lastline=195]{bst.v}

The correctness of the function is proved through lemma \texttt{treeMin\_is\_min} (line 519), which proves that for all trees, if the tree minimum is a natural number, i.e. the tree is not a leaf, then it occurs in the tree and it is less than or equal to all of its elements. To prove this lemma, we decided to split it into two trivial parts (sublemmas) due to the total length of the proofs and the required additional sublemmas for one of the parts:

\begin{itemize}
    \item \texttt{treeMin\_occurs}, which proves the first part, i.e. that if a tree has a natural number as minimum, then it occurs in the tree. The proof is quite long as we needed to prove for the different ramifications of the tree during the execution of \texttt{treeMin} and \texttt{occurs}.
    \lstinputlisting[firstnumber=198,firstline=198,lastline=318]{bst.v}
    
    \item \texttt{treeMin\_is\_min}, which proves the second part, i.e. that if a tree has a minimum, it is smaller or equal than all of the elements of the tree. It is hard to prove this sublemma as it is, so we want to prove first a looser version for any number that is less than or equal to the minimum.
    
    \lstinputlisting[firstnumber=510,firstline=510,lastline=516]{bst.v}
    
    Thus, to prove this sublemma, we further create a function and two additional sublemmas as described below, and in the end, the sublemma is a direct consequence of them:
    \begin{itemize}
        \item Function \texttt{helper\_ge}, which serves almost completely as a greater or equal check for two \texttt{option nat} values. It is only ever used with a tree minimum as first value, and its purpose is to imply that if the second value is some smaller or equal natural number, then all elements in the tree are greater than or equal to it. Therefore, when the second value is \texttt{None}, we return \texttt{False}, as the elements in the tree can't be compared to it, and otherwise, when the first value (tree minimum) is \texttt{None}, we return \texttt{True}, as there are no elements in the tree to compare to the second value. The case when they are both natural numbers is a regular greater or equal check.
        \lstinputlisting[firstnumber=322,firstline=322,lastline=332]{bst.v}
        
        \item \texttt{le\_than\_treeMin\_is\_le\_than\_all}, which proves that if a tree has no minimum (it is a leaf) or the minimum is greater than or equal to some natural number (using \texttt{helper\_ge}), then all values in the tree are greater than or equal to that number. This proof is also quite long as we needed to prove for the different ramifications of the tree during the execution of \texttt{treeMin} and \texttt{helper\_ge}.
        \lstinputlisting[firstnumber=336,firstline=336,lastline=442]{bst.v}
        
        \item \texttt{equal\_implies\_greater\_or\_equal}, which proves that if the tree minimum of a tree is equal to some natural number, then it is also greater than or equal to that number (in the sense of \texttt{helper\_ge}). As in the proof of the previous sublemma, there are quite some cases for \texttt{treeMin} and \texttt{helper\_ge} to be covered.
        \lstinputlisting[firstnumber=446,firstline=446,lastline=507]{bst.v}
    \end{itemize}
\end{itemize}

The proof of the lemma follows \texttt{treeMin\_correct}:

\lstinputlisting[firstnumber=519,firstline=519,lastline=526]{bst.v}
 
\paragraph{Leftmost Node}

The recursively defined leftmost function (\texttt{leftmost}) returns the value of the leftmost node of a tree (as \texttt{option nat}). In case the tree is a leaf, \texttt{None} is returned.

\lstinputlisting[firstnumber=529,firstline=529,lastline=536]{bst.v}

We then want to prove lemma \texttt{leftmost\_is\_min\_bst}, i.e. that for all BSTs, that the minimal element is the leftmost node. This proof is also quite long as we needed to prove for the different ramifications of the tree during the execution of \texttt{treeMin} and \texttt{leftmost}. Here the aforementioned sublemma \texttt{treeMin\_occurs} is quite useful, which combined with the sublemma \texttt{valid\_for\_all\_valid\_for\_one}, i.e. if a condition is valid for all values in a tree and a specific value occurs in that tree, then the condition is also valid for that value, give us some comparison between tree minimums of different subtrees, which helps us prove the different cases within the lemma.

\lstinputlisting[firstnumber=538,firstline=538,lastline=611]{bst.v}

\paragraph{Search a BST}

The recursively defined search function (\texttt{search}) searches a BST and checks if specific number occurs in it, leveraging the fact that it is a BST. Trivially, in case the tree is a leaf, \texttt{False} is returned.

\lstinputlisting[firstnumber=614,firstline=614,lastline=623]{bst.v}

We then want to prove lemma \texttt{search\_eq\_occurs}, i.e. that for all BSTs, \texttt{search} returns the same result as \texttt{occurs}. The proof of this lemma does not require any new sublemmas, but we do make use of the aforementioned sublemma \texttt{valid\_for\_all\_valid\_for\_one}. The proof is fairly straightforward, with induction on the tree in both directions, split into simple different cases for \texttt{search}.

\lstinputlisting[firstnumber=626,firstline=626,lastline=656]{bst.v}

\newpage
\appendix

\section{Code}

\lstinputlisting{bst.v}
\end{document}